---
title: "Analysis22"
author: "Chase Clark"
date: "January 19, 2019"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```


The IDBac version used to create the SQL was:  https://github.com/chasemc/IDBacApp/commit/9c7609ade28155e165c6e4319f12b55f6108b216

This can be installed via: 
devtools::install_github("chasemc/IDBacApp@9c7609a")

```{r message=FALSE, warning=FALSE}
library(RSQLite)
library(DBI)
library(dplyr)
library(magrittr)
library(MALDIquant)
```


Connect to SQLite, show available tables
```{r}
sqliteFile <- "data/r01_replicates.sqlite"


con <- DBI::dbConnect(drv = RSQLite::SQLite(),
                      sqliteFile)
DBI::dbListTables(con)
```

All raw and processed spectra, peaklists are contained in the table "IndividualSpectra" 
```{r}
spectraTable <- dplyr::tbl(con, "IndividualSpectra")
```




```{r}
spectraTable %>% 
  filter(!is.na(proteinPeaks)) %>% 
  select(Strain_ID, proteinPeaks) %>% 
  collect %>% 
  return(.) -> proteinPeaksTable

proteinPeaksTable %>% 
  mutate(Strain_ID, Strain = sapply(strsplit(Strain_ID, "_"), "[", 1)) %>% 
  mutate(Strain_ID, Bio = sapply(strsplit(Strain_ID, "_"), "[", 3)) %>%
  mutate(Strain_ID, Tech = sapply(strsplit(Strain_ID, "_"), "[", 5)) %>%
  filter(Strain %in% c("A019","A046", "B004", "172-6")) %>% 
  return(.) -> proteinPeaksTable


z<- unlist(lapply(proteinPeaksTable$proteinPeaks, function(x) unserialize(memDecompress(x, type= "gzip"))))

proteinPeaksTable %>% 
  mutate(proteinPeaks = z ) -> proteinPeaksTable
remove(z)



proteinPeaksTable$proteinPeaks <- lapply(proteinPeaksTable$proteinPeaks, function(x){
  
  x@intensity <- x@intensity / max(x@intensity)
  x
})



```



```{r echo=FALSE}

proteinPeaksTable$proteinPeaks %>% 
  MALDIquant::trim(., c(3000, 15000)) -> proteinPeaks



biologicalReplicates <- paste0(proteinPeaksTable$Strain, "_", proteinPeaksTable$Bio)

groupedByBio <- split(proteinPeaks, biologicalReplicates)


some <- lapply(groupedByBio, function(x){
  
  sample(x=x, 1)
})


ss <- split(some, sapply(strsplit(names(groupedByBio), "_"),"[",1))


sw <- lapply(ss, function(x){
  x <- unlist(x)
  z <- determineWarpingFunctions(x, tolerance=0.002,
                          method=c("lowess"),
                          allowNoMatches=FALSE,
                          plot=FALSE, plotInteractive=TRUE)
  x <-warpMassPeaks(x,z)
  x <- MALDIquant::binPeaks(unlist(x))
  x <- MALDIquant::filterPeaks(x, minFrequency = .5)
  MALDIquant::mergeMassPeaks(x)
  
})

ii <- MALDIquant::binPeaks(sw)

ii <- MALDIquant::intensityMatrix(ii)

ii[is.na(ii)] <- 0


coop::tcosine(ii)

```




## Biological variability.
Boxplots represent variability of technical-replicates cosine score within each of 12 biological replicates.

```{r echo=FALSE}




q3 <- lapply(grouped_by_strain, function(a){
q <- lapply(1:100, function(xx){
  
  a1 <-  lapply(a, function(x) sample(x, 1))
  
  a1 %>% 
    unlist %>%
    MALDIquant::trim(., c(3000, 15000)) %>% 
    MALDIquant::binPeaks(.) %>% 
    MALDIquant::intensityMatrix() -> a2
  
  a2[is.na(a2)] <- 0
  a3 <- coop::tcosine(a2)
  
})
return(do.call(rbind, q))
})


par(mfrow=c(2,2))

for(i in seq_along(q3)){
boxplot(q3[[i]], ylim= c(0,1), 
        main = names(q3)[[i]],
        xlab = "Biological Replicates",
        ylab = "Cosine Score")
}


         
  
```


```{r}

a <- grouped_by_strain[[2]]
# randomly sample one technical replicate within each biological replicate
a1 <-  lapply(a, function(x) sample(x, 1))


b <- grouped_by_strain[[3]]
# randomly sample one technical replicate within each biological replicate
b1 <-  lapply(b, function(x) sample(x, 1))


binned <- MALDIquant::binPeaks(unlist(list(a1,b1), recursive = T))

binned <- intensityMatrix(binned)

binned[is.na(binned)] <- 0

binned <- coop::tcosine(binned)

d <- binned
d[upper.tri(d, diag = T)] <- NA

a1 <- as.vector(d[1:12, 1:12])
a2 <- as.vector(d[13:24, 1:12])
a3 <- as.vector(d[13:24, 13:24])

a1 <- a1[!is.na(a1)]
a2 <- a2[!is.na(a2)]
a3 <- a3[!is.na(a3)]




```


```{r eval=FALSE, include=FALSE}
t.test(a2,a1)
t.test(a2,a3)

B      <- 1000
t.vect <- vector(length=B)
p.vect <- vector(length=B)

for(i in 1:B){
  boot.c <- sample(a1, size=3, replace=T)
  boot.p <- sample(a2, size=3, replace=T)
  ttest  <- t.test(boot.c, boot.p)
  t.vect[i] <- ttest$statistic
  p.vect[i] <- ttest$p.value
}

windows()
  qqplot(t.vect, distribution="t", df=42)

```



```{r eval=FALSE, include=FALSE}
ppmv <- do.call(rbind, ppmv)
colnames(ppmv)[[5]] <- "ppm"


ggplot(data = ppmv) + 
    geom_density(aes(x=value, y= ..density.., fill = groups), alpha=0.4) +
    coord_cartesian(xlim = c(0,1)) +
    scale_fill_discrete(name = "Cosine Similarity Between:") +
    ggtitle("Cosine similarity between Bacillus and Proteobacteria") +
    theme(plot.title = element_text(hjust = 0.5)) +
    xlab("Cosine Similarity") +
    ylab("Density") +
    facet_wrap(~`ppm`, scales ="free_y")
```


