---
title: "Analysis"
author: "Chase Clark"
date: "January 23, 2019"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
self_contained: true


---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = F, cache=TRUE)
```


The IDBac version used to create the SQL was:  https://github.com/chasemc/IDBacApp/commit/9c7609ade28155e165c6e4319f12b55f6108b216

This can be installed via: 
devtools::install_github("chasemc/IDBacApp@9c7609a")


Means and standard deviations were computed by 1,000 random-sample bootstraps.



```{r message=FALSE, warning=FALSE}
library(RSQLite)
library(DBI)
library(dplyr)
library(magrittr)
library(MALDIquant)
library(ggplot2)
library(gridExtra)
library(grid)
```

```{r}
set.seed(42)
```


```{r}
sqliteFile <- "data/r01_replicates.sqlite"
sqliteFile <- "data/k2_calibration.sqlite"

con <- DBI::dbConnect(drv = RSQLite::SQLite(),
                      sqliteFile)
DBI::dbListTables(con)
```

```{r}
spectraTable <- dplyr::tbl(con, "IndividualSpectra")
```



```{r}
spectraTable %>% 
  filter(!is.na(proteinSpectrum)) %>% 
  select(Strain_ID, proteinSpectrum) %>% 
  collect %>% 
  return(.) -> proteinSpectraTable

proteinSpectraTable %>% 
  mutate(Strain_ID, Strain = sapply(strsplit(Strain_ID, "_"), "[", 1)) %>% 
  mutate(Strain_ID, Bio = sapply(strsplit(Strain_ID, "_"), "[", 3)) %>%
  mutate(Strain_ID, Tech = sapply(strsplit(Strain_ID, "_"), "[", 5)) %>%
  filter(Strain %in% c("A019","A046", "B004", "172-6")) %>% 
  return(.) -> proteinSpectraTable


z<- unlist(lapply(proteinSpectraTable$proteinSpectrum, function(x) unserialize(memDecompress(x, type= "gzip"))))

proteinSpectraTable %>% 
  mutate(proteinSpectrum = z ) -> proteinSpectraTable




proteinSpectraTable$proteinSpectrum <- lapply(proteinSpectraTable$proteinSpectrum, function(x){
  
  x@intensity <- x@intensity / max(x@intensity) *100
  x
})





```



```{r echo=FALSE}
proteinSpectraTable$proteinSpectrum %>% 
  MALDIquant::smoothIntensity(.,
                              method = "SavitzkyGolay", 
                              halfWindowSize = 20) %>% 
  MALDIquant::removeBaseline(.,
                             method = "TopHat") %>% 
  MALDIquant::detectPeaks(., 
                          method = "MAD", 
                          halfWindowSize = 20, 
                          SNR = 6)  %>% 
  return(.) -> p


p %>% 
  MALDIquant::trim(., c(4000, 20000)) -> proteinPeaks



biologicalReplicates <- paste0(proteinSpectraTable$Strain, "_", proteinSpectraTable$Bio)

groupedByBio <- split(proteinPeaks, biologicalReplicates)

ids <- sapply(strsplit(names(groupedByBio), "_"), "[", 1)

  # Merge technical replicates

saveRDS(groupedByBio, "data/groupedByBio.rds")
saveRDS(ids, "data/ids.rds")


theFunc <- function(bio, tech, boot){

  techReps <- tech
  bioReps <- bio
  
  techSpec <- lapply(groupedByBio, 
                     function(x){
                       x <- sample(x, techReps)
                       x <- MALDIquant::binPeaks(x,
                                                 method = "strict",
                                                 tolerance = 0.002)
                       x <- MALDIquant::filterPeaks(x, 
                                                    minFrequency = 0.5)
                       MALDIquant::mergeMassPeaks(x,
                                                  method = "sum",
                                                  ignore.na = FALSE)
                     })


  
  
  bioSpec <- lapply(split(techSpec, ids), 
                     function(x){
                       
                       x <- sample(x, bioReps)
                  
                       x <- MALDIquant::binPeaks(x,
                                                 method = "relaxed",
                                                 tolerance = 0.002)
                       x <- MALDIquant::filterPeaks(x, 
                                                    minFrequency = 0.5)
                       MALDIquant::mergeMassPeaks(x,
                                                  method = "sum",
                                                  ignore.na = FALSE)
                     })






binned <- MALDIquant::binPeaks(bioSpec, tolerance = .002)

binned <- intensityMatrix(binned)

binned[is.na(binned)] <- 0

binned <- 1 - coop::tcosine(binned)
colnames(binned) <- names(bioSpec)
diag(binned) <- NA



return(cbind.data.frame(binned,
      techReps = techReps,
      bioReps = bioReps,
      boot = boot
      ))

}


```

```{r eval=FALSE, include=FALSE}


 numCores <- parallel::detectCores()
numCores <- parallel::makeCluster(numCores - 2) #Setup cluster
pp <- parallel::parLapply(numCores, 1:1000, function(boot) {
  set.seed(42)

library(MALDIquant)
  groupedByBio <- readRDS("data/groupedByBio.rds")
 ids <- readRDS("data/ids.rds")
  a<-expand.grid(1:3, 1:12)
  
theFunc <- function(bio, tech, boot){

  techReps <- tech
  bioReps <- bio
  
  techSpec <- lapply(groupedByBio, 
                     function(x){
                       x <- sample(x, techReps)
                       x <- MALDIquant::binPeaks(x,
                                                 method = "strict",
                                                 tolerance = 0.002)
                       x <- MALDIquant::filterPeaks(x, 
                                                    minFrequency = 0.5)
                       MALDIquant::mergeMassPeaks(x,
                                                  method = "sum",
                                                  ignore.na = FALSE)
                     })


  
  
  bioSpec <- lapply(split(techSpec, ids), 
                     function(x){
                       
                       x <- sample(x, bioReps)
                  
                       x <- MALDIquant::binPeaks(x,
                                                 method = "strict",
                                                 tolerance = 0.002)
                       x <- MALDIquant::filterPeaks(x, 
                                                    minFrequency = 0.5)
                       MALDIquant::mergeMassPeaks(x,
                                                  method = "sum",
                                                  ignore.na = FALSE)
                     })






binned <- MALDIquant::binPeaks(bioSpec, tolerance = .002)

binned <- intensityMatrix(binned)

binned[is.na(binned)] <- 0

binned <- 1 - coop::tcosine(binned)
colnames(binned) <- names(bioSpec)
diag(binned) <- NA



return(cbind.data.frame(binned,
      techReps = techReps,
      bioReps = bioReps,
      boot = boot
      ))

}


  mapply(theFunc, bio = a$Var2, tech = a$Var1, boot = boot,  SIMPLIFY = FALSE)

})



parallel::stopCluster(numCores)

saveRDS(pp, "data/1000-boots.rds")
```

```{r}
pp <- readRDS("data/1000-boots.rds")
zap <- do.call(rbind.data.frame, unlist(pp, recursive = F))

```

---

# Cosine Correlations 


## Comparing *Streptomyces*


```{r}
zap %>% 
    filter(!is.na(B004)) %>%
    filter(is.na(A019)) -> pl

zap %>% 
    filter(!is.na(B004)) %>%
    filter(is.na(A046)) -> pl2

pl$Group <-"B004 vs A019"

pl2$Group <-"B004 vs A046"
combined <- rbind(pl, pl2)

combined %>%
  mutate(cc = paste0(Group, "- ", techReps, " technical")) -> combined
    


ggplot(combined) + 
    geom_boxplot(aes(y=B004, x=factor(bioReps),fill=factor(cc))) +
    coord_cartesian(ylim=c(0,1)) + 
    labs(colour="Technical Replicates") +
    xlab("Biological Replicates") +
    ylab("Cosine Similarity") +
    ggtitle("The Effect of Biological and Technical Replicates On Discriminating Streptomyces") 

    
    
    
```


### Comparing *Streptomyces* to *Enterococcus*
```{r}
zap %>% 
    filter(!is.na(B004)) %>%
    filter(is.na(`172-6`)) -> pl

zap %>% 
    filter(!is.na(B004)) %>%
    filter(is.na(A046)) -> pl2

pl$Group <-"B004 vs 172-6"

pl2$Group <-"B004 vs A046"
combined <- rbind(pl, pl2)

combined %>%
  mutate(cc = paste0(Group, "- ", techReps, " technical")) -> combined
    


ggplot(combined) + 
    geom_boxplot(aes(y=B004, x=factor(bioReps),fill=factor(cc))) +
    coord_cartesian(ylim=c(0,1)) + 
    labs(colour="Technical Replicates") +
    xlab("Biological Replicates") +
    ylab("Cosine Similarity") +
    ggtitle("The Effect of Biological and Technical Replicates On Discriminating Streptomyces") 

    
    
    
```





```{r}
zap %>% 
    filter(!is.na(A046)) %>%
    filter(is.na(`172-6`)) -> pl

zap %>% 
    filter(!is.na(A046)) %>%
    filter(is.na(B004)) -> pl2

pl$Group <-"A046 vs 172-6"

pl2$Group <-"A046 vs B004"
combined <- rbind(pl, pl2)

combined %>%
  mutate(cc = paste0(Group, "- ", techReps, " technical")) -> combined
    


ggplot(combined) + 
    geom_boxplot(aes(y=A046, x=factor(bioReps),fill=factor(cc))) +
    coord_cartesian(ylim=c(0,1)) + 
    labs(colour="Technical Replicates") +
    xlab("Biological Replicates") +
    ylab("Cosine Similarity") +
    ggtitle("The Effect of Biological and Technical Replicates On Discriminating Streptomyces") 

    
    
    
```


# Boxplot Comparisons of Biological andf Technical Replicates 

This section shows the effects of 1 to 3 technical replicates at each of 12 biological replicates.

Technical replicates were used to create a merged peak list at each biological replicate, then each bioligcal replicate merged.
Peak-handling (eg binning settings) were kept the same for both technical and biological replicates. Percent peak presence was kept at 50% for both technical and biological replicates.


The boxplots below depict each of the four isolates against each other. 
As a reminder: 
- B004 and A046 share 100% 16S rRNA sequence
- A046 and A019 share 98.03% 16S rRNA sequence (29 nt)
- 172-6 is an *Enterococcus* sp.
See the Study Design  document for more details

## B004

### B004 vs A019

```{r}
zap %>% 
    filter(!is.na(B004)) %>%
    filter(is.na(A019)) %>% 
    
    ggplot() + 
    geom_boxplot(aes(y=B004, x=factor(bioReps),fill=factor(techReps)))+
    coord_cartesian(ylim=c(0,1))
```

### B004 vs A046

```{r}
zap %>% 
    filter(!is.na(B004)) %>%
    filter(is.na(A046)) %>% 
    
    ggplot() + 
    geom_boxplot(aes(y=B004, x=factor(bioReps),fill=factor(techReps)))+
    coord_cartesian(ylim=c(0,1))
```





### B004 vs 172-6


```{r}
zap %>% 
    filter(!is.na(B004)) %>%
    filter(is.na(`172-6`)) %>% 
    
    ggplot() + 
    geom_boxplot(aes(y=B004, x=factor(bioReps),fill=factor(techReps)))+
    coord_cartesian(ylim=c(0,1))
```


## A046


### A046 vs 172-6


```{r}
zap %>% 
    filter(!is.na(A046)) %>%
    filter(is.na(`172-6`)) %>% 
    
    ggplot() + 
    geom_boxplot(aes(y=A046, x=factor(bioReps), fill=factor(techReps)))+
    coord_cartesian(ylim=c(0,1))
```


### A046 vs A019


```{r}
zap %>% 
    filter(!is.na(A046)) %>%
    filter(is.na(A019)) %>% 
    
    ggplot() + 
    geom_boxplot(aes(y=A046, x=factor(bioReps), fill=factor(techReps)))+
    coord_cartesian(ylim=c(0,1))
```


### A046 vs B004


```{r}
zap %>% 
    filter(!is.na(A046)) %>%
    filter(is.na(B004)) %>% 
    
    ggplot() + 
    geom_boxplot(aes(y=A046, x=factor(bioReps), fill=factor(techReps)))+
    coord_cartesian(ylim=c(0,1))
```



## A019


### A019 vs 172-6


```{r}
zap %>% 
    filter(!is.na(A019)) %>%
    filter(is.na(`172-6`)) %>% 
    
    ggplot() + 
    geom_boxplot(aes(y=A019, x=factor(bioReps), fill=factor(techReps)))+
    coord_cartesian(ylim=c(0,1))
```


### A019 vs A046


```{r}
zap %>% 
    filter(!is.na(A019)) %>%
    filter(is.na(A046)) %>% 
    
    ggplot() + 
    geom_boxplot(aes(y=A019, x=factor(bioReps), fill=factor(techReps)))+
    coord_cartesian(ylim=c(0,1))
```


### A019 vs B004


```{r}
zap %>% 
    filter(!is.na(A019)) %>%
    filter(is.na(B004)) %>% 
    
    ggplot() + 
    geom_boxplot(aes(y=A019, x=factor(bioReps), fill=factor(techReps)))+
    coord_cartesian(ylim=c(0,1))
```






---
---


# Effect of Technical Replicates with increaing Biological Replicates (Standard Deviations)

This section of the document depicts the effect of adding between 1 and 3 technical Replicates in scenarios where 1 to 12 biological replicates are included. The measurements depicted are the standard deviation of cosine similarity.

*Note: It is important to remember here that at each biological replicate level, technical replicates are not additive but **multiply** the number of samples*



```{r}
sub <- function(x){ x[[1]] - x[[2]]}

data <- dplyr::as.tbl(zap)
plotter <- function(data, one, two){

  data %>% 
    filter(!is.na(UQ(sym(one)))) %>%
    filter(is.na(UQ(sym(two)))) %>% 
    group_by(boot, techReps,bioReps) %>% 
    summarise(mean = mean(UQ(sym(one)))) %>% 
    ungroup %>% 
    group_by(techReps, bioReps) %>% 
        summarise(sd= sd(mean), mean = mean(mean)) %>% 
   
     ggplot(data = .) + 
  geom_point(aes(x = factor(bioReps),
                 y = sd)) +
  geom_line(aes(x = bioReps,
                y = sd,
                group = factor(techReps),
                color = factor(techReps))) +
  labs(colour="Technical Replicates") +
  xlab("Biological Replicates") +
  ylab("Standard Deviation of Cosine Similarity") +
  ggtitle(paste0("Effect of Technical Replication with Increasing Biological Replicates \n ",
                one, " vs ", two)) +
    coord_cartesian(ylim=c(0,.2))%>% 
    return(.) 
}
  



```

## Effects measured against A046

```{r,  fig.width=12, fig.height=12}
gridExtra::grid.arrange(grid::textGrob(""),
                        plotter(data, "A046","B004"),
                        plotter(data, "A046","A019"),
                        plotter(data, "A046","172-6"),
                        ncol=2, nrow=2)
```

## Effects measured against B004

```{r,  fig.width=12, fig.height=12}
gridExtra::grid.arrange(plotter(data, "B004","A046"),
                        grid::textGrob(""),
                        plotter(data, "B004","A019"),
                        plotter(data, "B004","172-6"),
                        ncol=2, nrow=2)
```

## Effects measured against A019

```{r,  fig.width=12, fig.height=12}
gridExtra::grid.arrange(plotter(data, "A019","A046"),
                        plotter(data, "A019","B004"),
                        grid::textGrob(""),
                        plotter(data, "A019","172-6"),
                        ncol=2, nrow=2)
```

## Effects measured against 172-6

```{r,  fig.width=12, fig.height=12}
gridExtra::grid.arrange(plotter(data, "172-6","A046"),
                        plotter(data, "172-6","B004"),
                        plotter(data, "172-6","A019"),
                        grid::textGrob(""),
                        ncol=2, nrow=2)
```





---
---

# Effect of Technical Replicates with increaing Biological Replicates (Means)

This section of the document depicts the effect of adding between 1 and 3 technical Replicates in scenarios where 1 to 12 biological replicates are included. The measurements depicted are the means of cosine similarity.

*Note: It is important to remember here that at each biological replicate level, technical replicates are not additive but **multiply** the number of samples*


```{r}
sub <- function(x){ x[[1]] - x[[2]]}

data <- dplyr::as.tbl(zap)
plotter <- function(data, one, two){

  data %>% 
    filter(!is.na(UQ(sym(one)))) %>%
    filter(is.na(UQ(sym(two)))) %>% 
    group_by(boot, techReps,bioReps) %>% 
    summarise(mean = mean(UQ(sym(one)))) %>% 
    ungroup %>% 
    group_by(techReps, bioReps) %>% 
        summarise(sd= sd(mean), mean = mean(mean)) %>% 
   
     ggplot(data = .) + 
  geom_point(aes(x = factor(bioReps),
                 y = mean)) +
  geom_line(aes(x = bioReps,
                y = mean,
                group = factor(techReps),
                color = factor(techReps))) +
  labs(colour="Technical Replicates") +
  xlab("Biological Replicates") +
  ylab("Standard Deviation of Cosine Similarity") +
  ggtitle(paste0("Effect of Technical Replication with Increasing Biological Replicates \n ",
                one, " vs ", two)) +
    coord_cartesian(ylim=c(0,1))%>% 
    return(.) 
}
  



```


## Effects measured against A046

```{r,  fig.width=12, fig.height=12}
gridExtra::grid.arrange(grid::textGrob(""),
                        plotter(data, "A046","B004"),
                        plotter(data, "A046","A019"),
                        plotter(data, "A046","172-6"),
                        ncol=2, nrow=2)
```

## Effects measured against B004


```{r,  fig.width=12, fig.height=12}
gridExtra::grid.arrange(plotter(data, "B004","A046"),
                        grid::textGrob(""),
                        plotter(data, "B004","A019"),
                        plotter(data, "B004","172-6"),
                        ncol=2, nrow=2)
```

## Effects measured against A019


```{r,  fig.width=12, fig.height=12}
gridExtra::grid.arrange(plotter(data, "A019","A046"),
                        plotter(data, "A019","B004"),
                        grid::textGrob(""),
                        plotter(data, "A019","172-6"),
                        ncol=2, nrow=2)
```

## Effects measured against 172-6


```{r,  fig.width=12, fig.height=12}
gridExtra::grid.arrange(plotter(data, "172-6","A046"),
                        plotter(data, "172-6","B004"),
                        plotter(data, "172-6","A019"),
                        grid::textGrob(""),
                        ncol=2, nrow=2)
```







----

# Evaluating the effects of increasing MALDI sample spots


This section 

In this case it looks like increasing sample spots, in general, is the most effect strategy. 






```{r}
sub <- function(x){ x[[1]] - x[[2]]}

plotter <- function(data, one, two){

data %>% 
  filter(!is.na(UQ(sym(one)))) %>%
  filter(is.na(UQ(sym(two)))) %>% 
  group_by(boot, techReps,bioReps) %>% 
  summarise(mean = mean(UQ(sym(one)))) %>% 
  ungroup %>% 
  group_by(techReps, bioReps) %>% 
  summarise(sd= sd(mean), mean = mean(mean)) %>% 
  mutate(spots = bioReps * techReps) -> pp

p <- ggplot(pp) +
  geom_point(aes(x=spots, y=sd, color=factor(techReps)))+
  geom_line(aes(x=spots, y=sd, color=factor(techReps))) +
  labs(colour="Technical Replicates") +
  xlab("MALDI Sample Spots") +
  ylab("Standard Deviation of Cosine Similarity") +
  ggtitle(paste0("Effect of Technical Replication with Increasing Biological Replicates \n ",
                 one, " vs ", two)) 

p2 <- ggplot(pp) +
  geom_point(aes(x=spots, y=mean, color=factor(techReps))) +
  geom_line(aes(x=spots, y=mean, color=factor(techReps))) +
  labs(colour="Technical Replicates") +
  xlab("MALDI Sample Spots") +
  ylab("Mean of Cosine Similarity") +
  ggtitle(paste0("Effect of Technical Replication with Increasing Biological Replicates \n ",
                 one, " vs ", two)) 

return(list(p,
            p2))
}
  
```


```{r,  fig.width=12, fig.height=12}
a1 <- plotter(data, "A046","A019")
a2 <- plotter(data, "A046","B004")
a3 <- plotter(data, "A046","172-6")

a1 <- gridExtra::arrangeGrob(a1[[1]], a1[[2]], ncol = 2)
a2 <- gridExtra::arrangeGrob(a2[[1]], a2[[2]], ncol = 2)
a3 <- gridExtra::arrangeGrob(a3[[1]], a3[[2]], ncol = 2)

z<-gridExtra::arrangeGrob(a1,a2,a3, ncol=1)
grid::grid.draw(z)
```


```{r,  fig.width=12, fig.height=12}
a1 <- plotter(data, "A019","A046")
a2 <- plotter(data, "A019","B004")
a3 <- plotter(data, "A019","172-6")

a1 <- gridExtra::arrangeGrob(a1[[1]], a1[[2]], ncol = 2)
a2 <- gridExtra::arrangeGrob(a2[[1]], a2[[2]], ncol = 2)
a3 <- gridExtra::arrangeGrob(a3[[1]], a3[[2]], ncol = 2)

z<-gridExtra::arrangeGrob(a1,a2,a3, ncol=1)
grid::grid.draw(z)
```


```{r,  fig.width=12, fig.height=12}
a1 <- plotter(data, "B004","A046")
a2 <- plotter(data, "B004","A019")
a3 <- plotter(data, "B004","172-6")

a1 <- gridExtra::arrangeGrob(a1[[1]], a1[[2]], ncol = 2)
a2 <- gridExtra::arrangeGrob(a2[[1]], a2[[2]], ncol = 2)
a3 <- gridExtra::arrangeGrob(a3[[1]], a3[[2]], ncol = 2)

z<-gridExtra::arrangeGrob(a1,a2,a3, ncol=1)
grid::grid.draw(z)
```


```{r,  fig.width=12, fig.height=12}
a1 <- plotter(data, "172-6","A046")
a2 <- plotter(data, "172-6","A019")
a3 <- plotter(data, "172-6","B004")

a1 <- gridExtra::arrangeGrob(a1[[1]], a1[[2]], ncol = 2)
a2 <- gridExtra::arrangeGrob(a2[[1]], a2[[2]], ncol = 2)
a3 <- gridExtra::arrangeGrob(a3[[1]], a3[[2]], ncol = 2)

z<-gridExtra::arrangeGrob(a1,a2,a3, ncol=1)

grid::grid.draw(z)
```

