---
title: "Analysis3"
author: "Chase Clark"
date: "January 23, 2019"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
self_contained: true
---

```{r setup, include=FALSE, }
knitr::opts_chunk$set(echo = F, cache=TRUE)
```


The IDBac version used to create the SQL was:  https://github.com/chasemc/IDBacApp/commit/9c7609ade28155e165c6e4319f12b55f6108b216

This can be installed via: 
devtools::install_github("chasemc/IDBacApp@9c7609a")


Means and standard deviations were computed by 1,000 random-sample bootstraps.



```{r message=FALSE, warning=FALSE}
library(RSQLite)
library(DBI)
library(dplyr)
library(magrittr)
library(MALDIquant)
library(ggplot2)
library(gridExtra)
library(grid)
```

```{r}
set.seed(42)
```


```{r}
con <- IDBacApp::createPool(fileName = "replicate_study", 
                            filePath = here::here("data",
                                                  "sqlite"))[[1]]

DBI::dbListTables(con)
```


Get the spcetra data
```{r}
spectraTable <- dplyr::tbl(con, "IndividualSpectra") %>% collect()
```




Get all peak data. In the the sample map technical replicates share the same name as the biological replicate.  That means  `IDBacApp::getPeakData` will return a list of lists of MALDIquant objects, where each element of the top list is a biological replicate that contains a list of technical replicate MALDIquant objects. Because this is confusing I will show the first two elements of the output...

`172-6_1` and `172-6_10` are biological replicates `1` and `10` of strain `172-6`. Each of these two biological replicates contains four technical replicates.

```{r}
peak_data <- IDBacApp::getPeakData(con, 
                       sampleIDs = unique(spectraTable$Strain_ID),
                     protein = T) 


peak_data <- lapply(peak_data, function(x) MALDIquant::trim(x, c(4000, 20000)))

peak_data[1:2]
```


Not a good thing to do but a single biological replicate is missing a sample, here one will be copied as a fourth

```{r}
peak_data[[which(lengths(peak_data) == 3)]] <- c(peak_data[which(lengths(peak_data) == 3)], peak_data[which(lengths(peak_data) == 3)][[1]])
```


```{r}
lapply(peak_data, function(x){
  
  x[]
  
  
})
```

ghkvfthghvjhvjhvjhvjhvjhvjhjkdryh

```{r echo=FALSE}

theFunc <- function(bio, tech, boot){
  
  techSpec <- lapply(peak_data, 
                     function(x){
                       x <- MALDIquant::binPeaks(x,
                                                 method = "strict",
                                                 tolerance = 0.002)
                       x <- MALDIquant::filterPeaks(x, 
                                                    minFrequency = 0.5)
                       MALDIquant::mergeMassPeaks(x,
                                                  method = "sum",
                                                  ignore.na = FALSE)
                     })
  
  
  
  
  bioSpec <- lapply(split(techSpec, ids), 
                    function(x){
                      
                      x <- sample(x, bioReps)
                      
                      x <- MALDIquant::binPeaks(x,
                                                method = "relaxed",
                                                tolerance = 0.002)
                      x <- MALDIquant::filterPeaks(x, 
                                                   minFrequency = 0.5)
                      MALDIquant::mergeMassPeaks(x,
                                                 method = "sum",
                                                 ignore.na = FALSE)
                    })
  
  
  
  
  
  
  binned <- MALDIquant::binPeaks(bioSpec, tolerance = .002)
  
  binned <- intensityMatrix(binned)
  
  binned[is.na(binned)] <- 0
  
  binned <- 1 - coop::tcosine(binned)
  colnames(binned) <- names(bioSpec)
  diag(binned) <- NA
  
  
  
  return(cbind.data.frame(binned,
                          techReps = techReps,
                          bioReps = bioReps,
                          boot = boot
  ))
  
}


```

```{r eval=FALSE, include=FALSE}


numCores <- parallel::detectCores()
numCores <- parallel::makeCluster(numCores - 2) #Setup cluster
pp <- parallel::parLapply(numCores, 1:1000, 
                          function(boot) {
                            set.seed(42)
                            
                            library(MALDIquant)
                            groupedByBio <- readRDS("data/groupedByBio.rds")
                            ids <- readRDS("data/ids.rds")
                            a<-expand.grid(1:3, 1:12)
                            
                         
                            
                            
                            mapply(theFunc, bio = a$Var2, tech = a$Var1, boot = boot,  SIMPLIFY = FALSE)
                            
                          })



parallel::stopCluster(numCores)

saveRDS(pp, "data/1000-boots.rds")
```

```{r}
pp <- readRDS("data/1000-boots.rds")
zap <- do.call(rbind.data.frame, unlist(pp, recursive = F))

```



# Boxplot Comparisons of Biological andf Technical Replicates 

This section shows the effects of 1 to 3 technical replicates at each of 12 biological replicates.

Technical replicates were used to create a merged peak list at each biological replicate, then each bioligcal replicate merged.
Peak-handling (eg binning settings) were kept the same for both technical and biological replicates. Percent peak presence was kept at 50% for both technical and biological replicates.


The boxplots below depict each of the four isolates against each other. 
As a reminder: 
- B004 and A046 share 100% 16S rRNA sequence
- A046 and A019 share 98.03% 16S rRNA sequence (29 nt)
- 172-6 is an *Enterococcus* sp.
See the Study Design  document for more details

## B004

### B004 vs A019

```{r}
zap %>% 
  filter(!is.na(B004)) %>%
  filter(is.na(A019)) -> pl

zap %>% 
  filter(!is.na(B004)) %>%
  filter(is.na(A046)) -> pl2

pl$Group <-"B004 vs A019"

pl2$Group <-"B004 vs A046"
combined <- rbind(pl, pl2)

combined %>%
  mutate(cc = paste0(Group, "- ", techReps, " technical")) -> combined



ggplot(combined) + 
  geom_boxplot(aes(y=B004, x=factor(bioReps),fill=factor(cc))) +
  coord_cartesian(ylim=c(0,1)) + 
  labs(colour="Technical Replicates") +
  xlab("Biological Replicates") +
  ylab("Cosine Similarity") +
  ggtitle(paste0("Effect of Technical Replication with Increasing Biological Replicates \n ",
                 one, " vs ", two)) +
  coord_cartesian(ylim=c(0,.2))




```

### B004 vs A046

```{r}
zap %>% 
  filter(!is.na(B004)) %>%
  filter(is.na(A046)) %>% 
  
  
  
  ggplot() + 
  geom_boxplot(aes(y=B004, x=factor(bioReps),fill=factor(techReps)))+
  coord_cartesian(ylim=c(0,1))
```





### B004 vs 172-6


```{r}
zap %>% 
  filter(!is.na(B004)) %>%
  filter(is.na(`172-6`)) %>% 
  
  ggplot() + 
  geom_boxplot(aes(y=B004, x=factor(bioReps),fill=factor(techReps)))+
  coord_cartesian(ylim=c(0,1))
```


## A046


### A046 vs 172-6


```{r}
zap %>% 
  filter(!is.na(A046)) %>%
  filter(is.na(`172-6`)) %>% 
  
  ggplot() + 
  geom_boxplot(aes(y=A046, x=factor(bioReps), fill=factor(techReps)))+
  coord_cartesian(ylim=c(0,1))
```


### A046 vs A019


```{r}
zap %>% 
  filter(!is.na(A046)) %>%
  filter(is.na(A019)) %>% 
  
  ggplot() + 
  geom_boxplot(aes(y=A046, x=factor(bioReps), fill=factor(techReps)))+
  coord_cartesian(ylim=c(0,1))
```


### A046 vs B004


```{r}
zap %>% 
  filter(!is.na(A046)) %>%
  filter(is.na(B004)) %>% 
  
  ggplot() + 
  geom_boxplot(aes(y=A046, x=factor(bioReps), fill=factor(techReps)))+
  coord_cartesian(ylim=c(0,1))
```



## A019


### A019 vs 172-6


```{r}
zap %>% 
  filter(!is.na(A019)) %>%
  filter(is.na(`172-6`)) %>% 
  
  ggplot() + 
  geom_boxplot(aes(y=A019, x=factor(bioReps), fill=factor(techReps)))+
  coord_cartesian(ylim=c(0,1))
```


### A019 vs A046


```{r}
zap %>% 
  filter(!is.na(A019)) %>%
  filter(is.na(A046)) %>% 
  
  ggplot() + 
  geom_boxplot(aes(y=A019, x=factor(bioReps), fill=factor(techReps)))+
  coord_cartesian(ylim=c(0,1))
```


### A019 vs B004


```{r}
zap %>% 
  filter(!is.na(A019)) %>%
  filter(is.na(B004)) %>% 
  
  ggplot() + 
  geom_boxplot(aes(y=A019, x=factor(bioReps), fill=factor(techReps)))+
  coord_cartesian(ylim=c(0,1))
```




